A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-07-04 paper/fascd.tex
Some more edits

\usepackage{times,bm,bbm,empheq,verbatim,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{booktabs,multirow,xspace}

@@

We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).  Suppose $\Omega \subset \RR^d$ is a (open) bounded polygon.\footnote{Section \ref{sec:results} illustrates $d=1,2,3$ cases.}  We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a mesh hierarchy; the goal is an accurate solution on the finest-level mesh.  The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.

The mesh hierarchy and FE spaces can be constructed by standard nested refinement.  Suppose $\mathcal{T}^0$ is the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate elements with union equal to $\overline{\Omega}$ and maximum edge length (characteristic mesh size) $h_0>0$.  We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ element becomes $2^d$ elements in $\mathcal{T}^j$ with maximum edge length $h_j = 2^{-j} h_0$.  We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.  Let $m_j$ be the number of nodes in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.  Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, and observe the following vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}
On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.

In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
Also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions.  This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.

Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.  Note that the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.  We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
\begin{equation}
\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
\end{equation}
If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then, as for \eqref{eq:vi}, FE VI \eqref{eq:fe:vi} is well-posed.

@@

The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for FE VI \eqref{eq:fe:vi}, requires transfer operators between meshes.  We require five such operators, of which three are standard \cite{Trottenbergetal2001}.  Table \ref{tab:transfers} will help the reader keep track.

\begin{table}[H]
\begin{tabular}{llc}
\emph{name}  & \emph{action}  & \emph{linear?} \\ \hline
canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
(nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \\
minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ &
\end{tabular}

\medskip
\caption{Transfer operators used in this paper.}
\label{tab:transfers}
\end{table}

Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ is the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.  Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.  These canonical transfers generally involve a change of FE representation.  For example, $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.

@@

The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh hat functions.  To make this precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding hat functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so it follows that
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.

@@

Our practical multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.  The next Example concretely illustrates why we do not apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.  Our straightforward (iterate-dependent) bounds on the defect constraints, namely \eqref{eq:fe:chiordering} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.

\begin{example}  \label{ex:directRbad}  
For concreteness, suppose $\Omega = [0,L] \subset \RR^1$ is an interval and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the range of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.  Supposing empty Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematical.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels}---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty should be clear: the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

\begin{figure}[ht]
\begin{center}
\input{tikz/directRbad.tex}
\end{center}
\caption{For finest-level box constraints $\underline{\gamma}^1,\overline{\gamma}^1$ like these, direct restriction using any of $\iR,\minR,\maxR$, to a coarser mesh (solid dots), is problematic for admissibility of coarse corrections.}
\label{fig:directRbad}
\end{figure}

@@

This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are actually not required.

We now compute the LDC differences \cite{GraeserKornhuber2009}:
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute coarse corrections in $\mathcal{D}^j$ for the downward side and in $\mathcal{U}^j$ for the upward side.  This strategy is, however, based on \emph{incomplete} CDs, in the general box-constrained cases, as we now explain.  Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~\eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.  The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.

@@

\begin{example}  \label{ex:notfullcd}
Let $a\ge 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
\end{example}

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel FE schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.  Here there is, in fact, another ``basic'' iteration, namely \pr{icd-tele} in Algorithm \ref{alg:basiccd-tele}, which also does not need the $\Pi_i$ operators.  It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).

@@

While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs which are not of minimization type, and/or to box constraints, would seem to require full CDs.  We cannot offer that analysis.

@@

A key idea is that, because of inclusions \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.  Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus they must be in sufficiently small sets ($\mathcal{D}^j$) so that their sum is admissible once we return to a given level.  Going upward, however, no further coarse corrections can violate admissibility.

>>>>

\usepackage{times,bm,bbm,empheq,verbatim,fancyvrb,graphicx,amsthm,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{booktabs,multirow,xspace}
\usepackage{pifont}

@@

We seek to rapidly compute the approximate solution of box-constrained VI problems \eqref{eq:boxdirichletvi} using finite elements (FE).  Let $\Omega \subset \RR^d$ be a (open) bounded polygon. We will construct a multilevel FE approximation to \eqref{eq:boxdirichletvi} over $\Omega$ using $P_1$ or $Q_1$ elements \cite{Elmanetal2014}, by refining a coarsest mesh into a nested mesh hierarchy; the goal is an accurate solution on the finest-level mesh.  The hierarchy allows the creation of a multilevel CD of $\cK$, using the ideas in the next two Sections, which permits solvers based on admissibility-preserving coarse corrections.

The mesh hierarchy and FE spaces can be constructed by standard nested refinement.  Let $\mathcal{T}^0$ be the \emph{coarsest mesh}, a finite set of non-overlapping, nondegenerate cells with union equal to $\overline{\Omega}$ and maximum cell diameter (characteristic mesh size) $h_0>0$.  We assume $\mathcal{T}^0$ is conforming, with no hanging nodes.  For $j=1,\dots,J$, let $\mathcal{T}^j$ be the uniform refinement of $\mathcal{T}^{j-1}$ so that each $\mathcal{T}^{j-1}$ cell becomes $2^d$ cells in $\mathcal{T}^j$ with mesh size $h_j = 2^{-j} h_0$.  We call $J\ge 0$ the \emph{depth} of the hierarchy, $J+1$ the \emph{number of levels}, and $\mathcal{T}^J$ the \emph{finest-level mesh}.  Let $m_j$ be the number of vertices in $\mathcal{T}^j$, so there are $m_j$ \emph{degrees of freedom} at the $j$th level.  Let $\mathcal{V}^j$ be the continuous $P_1$ or $Q_1$ FE space over $\mathcal{T}^j$, with arbitrary boundary values, with the standard vector space nesting:
\begin{equation}
\mathcal{V}^0 \subset \mathcal{V}^1 \subset \dots \subset \mathcal{V}^J \subset \mathcal{V}=W^{1,p}(\Omega).  \label{eq:fe:nestedspaces}
\end{equation}
On the finest-level mesh let $g_D^J \in \mathcal{V}^J$ satisfy $g_D^J = g_D$ on $\partial_D \Omega$; this restricts $g_D$ in \eqref{eq:originalconstraintset} to be piecewise-linear.

In the material which follows, use of higher-order Lagrange elements, i.e.~$P_k$ or $Q_k$ with $k\ge 2$, would imply a ``variational crime'' \cite[Chapter 10]{BrennerScott2007} special to inequality constraints, which we avoid.  Namely, our FE construction of constraint sets (Section \ref{sec:cdmultilevel}) will require \emph{nodal monotonicity},
\begin{equation}
\bw \ge \bz \quad \implies \quad w \ge z \label{eq:nodalmonotonicity}
\end{equation}
for all $w,z \in \mathcal{V}^j$, where $\bw,\bz \in \RR^{m_j}$ denote vectors of nodal values.  However, for $k\ge 2$ there are hat functions in $P_k$ and $Q_k$, i.e.~functions with nodal values one or zero, which take on negative values \cite[Figure 1.7]{Elmanetal2014}.  Evidently \eqref{eq:nodalmonotonicity} is equivalent to the non-existence of such indefinite-sign hat functions.
\pef{I'm not sure this constitutes a variational crime? We enforce the RHS because it's tractable, and this is only equivalent to what we actually want (the LHS) for P1 and Q1. I don't see where we're departing from the variational framework, however.}

To allow box constraints on all meshes in the hierarchy, let $\tilde{\mathcal{V}}^j$ denote the set of functions on the nodes of $\mathcal{T}^j$ with values in the extended real line $\tilde{\RR} = [-\infty,+\infty]$.  Suppose that the finest-level \emph{lower} and \emph{upper obstacles} $\underline{\gamma}^J, \overline{\gamma}^J \in \tilde{\mathcal{V}}^J$ satisfy
\begin{equation}
\underline{\gamma}^J \le \overline{\gamma}^J, \qquad \underline{\gamma}^J < +\infty, \qquad -\infty < \overline{\gamma}^J. \label{eq:fe:boxconstraintrequirements}
\end{equation}
\pef{Shouldn't there be some relation between the continuous bounds and the discrete ones?}
In analogy to the continuous data, also assume that $\underline{\gamma}^J \le g_D^J \le \overline{\gamma}^J$ on those nodes of $\mathcal{T}^J$ which lie in $\partial_D \Omega$.  Define
\begin{equation}
\mathcal{K}^J = \left\{w\,:\,\underline{\gamma}^J \le w \le \overline{\gamma}^J \text{ on nodes of } \mathcal{T}^J, \, \text{ and } \, w|_{\partial_D\Omega} = g_D^J|_{\partial_D\Omega}\right\} \subset \mathcal{V}^J, \label{eq:fe:fineconstraintset}
\end{equation}
the finest-level \emph{constraint set} of admissible FE functions. This is a nonempty, closed, and convex subset of $\mathcal{V}^J$.  %As $\mathcal{V}^J$ functions never take values $\pm\infty$, inequality in \eqref{eq:fe:fineconstraintset} is strict where $\underline{\gamma}^J, \overline{\gamma}^J$ are infinite; there is no lower or upper constraint, respectively, at such nodes.

Let $f^J:\mathcal{K}^J \to (\mathcal{V}^J)'$ be a discretization of $f$ in \eqref{eq:boxdirichletvi}.  Note that the functional $\ell^J=\ell$ is defined and continuous over $\mathcal{V}^J$.  We seek the FE solution $u^J \in \mathcal{K}^J$ of the finest-level, finite-dimensional, nonlinear VI
\begin{equation}
\ip{f^J(u^J)}{v-u^J} \ge \ip{\ell^J}{v-u^J} \qquad \text{for all } v\in \cK^J. \label{eq:fe:vi}
\end{equation}
If $f^J$ is continuous and $p$-coercive on $\mathcal{K}^J$, in the sense that \eqref{eq:pcoercive} holds for all $u,v \in \mathcal{K}^J$, then the discrete VI \eqref{eq:fe:vi} is well-posed.

@@

The multilevel algorithm defined in Section \ref{sec:vcycle}, a solver for \eqref{eq:fe:vi}, requires five transfer operators between meshes.  Of these five operators, three are employed in standard FAS \cite{Trottenbergetal2001}.  We include Table \ref{tab:transfers} to help the reader keep track.

\begin{table}[H]
\begin{tabular}{llc}
\toprule
\emph{name}  & \emph{action}  & \emph{linear?} \\ \midrule
canonical prolongation        & $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ & \,\checkmark \\
canonical (dual) restriction  & $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ & \,\checkmark \\
(nodal) injection             & $\iR:\mathcal{V}^j\to\mathcal{V}^{j-1}$ & \,\checkmark \\
maximum restriction           & $\maxR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
minimum restriction           & $\minR:\tilde{\mathcal{V}}^j\to\tilde{\mathcal{V}}^{j-1}$ & \ding{55} \\
\bottomrule
\end{tabular}
\caption{Transfer operators used in this paper.}
\label{tab:transfers}
\end{table}

Canonical prolongation $P:\mathcal{V}^{j-1}\to\mathcal{V}^j$ implements the vector space inclusion $\mathcal{V}^{j-1} \hookrightarrow \mathcal{V}^j$ of coarser-mesh functions.  Canonical (dual) restriction $R:(\mathcal{V}^j)'\to(\mathcal{V}^{j-1})'$ also acts as an identity in the sense that if $\sigma \in (\mathcal{V}^j)'$ then $(R\sigma)[z] = \sigma[z]$ for all $z \in \mathcal{V}^{j-1} \subset \mathcal{V}^j$.  These canonical transfers generally involve a change of FE representation.  For example, $R$ is ``full-weighting'' \cite{Trottenbergetal2001}.

@@

\pef{I don't like the name ``monotone restrictions''. The word restriction is strongly associated with operating on the dual space. I would prefer to call them ``monotone injections'', since that is the matching operator.}
The monotone restrictions $\maxR,\minR$ \cite{GraeserKornhuber2009}, which will only be applied to nodal functions $\tilde{\mathcal{V}}^j$, are nonlinear maps which maximize and minimize, respectively, nodal values over the open supports of coarser-mesh basis functions (i.e.~over the \emph{stars} of the coarse vertices). These operators are constructed to enforce the property
\begin{equation}
\minR z \le z \le \maxR z,  \label{eq:fe:monotonerestrictionprops}
\end{equation}
which does not hold for the standard injection operator $\iR$.
To make the construction precise, let $x_q^j$ denote the nodes of mesh $\mathcal{T}^j$, with corresponding basis functions $\psi_q^j(x)$.  For a $\tilde\RR$-valued nodal function $z\in\tilde{\mathcal{V}}^j$ we define functions in $\tilde{\mathcal{V}}^{j-1}$:
\begin{align}
(\maxR z)(x_p^{j-1}) &= \max \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}, \label{eq:fe:monotonerestrictions} \\
(\minR z)(x_p^{j-1}) &= \min \{z(x_q^j) \,:\, \psi_p^{j-1}(x_q^j) > 0\}. \notag
\end{align}
For standard refinement a coarse-mesh node $x_p^{j-1}$ is in the set of fine nodes $x_q^j$ such that $\psi_p^{j-1}(x_q^j)>0$, so \eqref{eq:fe:monotonerestrictionprops} follows.
On the other hand, $z\ge 0$ implies $\minR z \ge 0$ and $z \le 0$ implies $\maxR z \le 0$.  These ordering properties are used in the CD construction of the next two Sections.
\pef{It would be good to have a diagram here. I could write one in tikz, if you don't have one handy.}

@@

A key choice in devising a multilevel scheme for VIs is how to construct coarse-grid bounds.
Our multilevel CD scheme uses \emph{defect constraints} \cite{GraeserKornhuber2009} which are based on an admissible finest-level iterate.  The next Example concretely illustrates why we do not apply restriction operators directly\footnote{An exception is in the F-cycle; see Section \ref{sec:implementation}.} to the finest-level box constraints $\underline{\gamma}^J,\overline{\gamma}^J$.  Our straightforward (iterate-dependent) bounds on the defect constraints, namely \eqref{eq:fe:chiordering} below, bypass the difficulties exposed in Example \ref{ex:directRbad}.

\begin{example}  \label{ex:directRbad}

Let $\Omega = (0,L) \subset \RR^1$ and consider a two-level mesh hierarchy ($J=1$).  Let $\underline{\gamma}^1$ be a sawtooth function with minimum value zero, and suppose $\overline{\gamma}^1=C+\underline{\gamma}^1$ with $C>0$ which is strictly less than the maximum of $\underline{\gamma}^1$.  Figure \ref{fig:directRbad} shows such functions; here levels $\mathcal{T}^1$ and $\mathcal{T}^0$ have $5$ and $3$ nodes, respectively.  Ignoring Dirichlet boundary for simplicity, the constraint set $\mathcal{K}^1$ defined by \eqref{eq:fe:fineconstraintset} is nonempty.  However, various restriction schemes applied directly to $\underline{\gamma}^1,\overline{\gamma}^1$ are problematic.  Suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \iR \underline{\gamma}^1, \iR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
Then $\underline{\gamma}^0=0$ and $\overline{\gamma}^0=C$ identically.  In that case the corresponding constraint set $\mathcal{K}^0$ is nonempty, but prolongation $Pw^0$ of an admissible solution of the coarse problem ($\underline{\gamma}^0 \le w^0 \le \overline{\gamma}^0$) is not admissible on the finer level: $Pw^0 \notin \mathcal{K}^1$.  On the other hand, suppose we propose
    $$\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1. \qquad \text{\emph{[problematic]}}$$
The general motivation for this choice---compare \eqref{eq:fe:chilevels} below---is that $w^0 \ge \underline{\gamma}^0 \, \implies \, Pw^0 \ge \underline{\gamma}^1$ and $w^0 \le \overline{\gamma}^0 \, \implies \, Pw^0 \le \overline{\gamma}^1$.  However, while these facts are true, the difficulty with this choice is that the coarse-level admissible set $\mathcal{K}^0$ is now empty because the proposed lower bound $\underline{\gamma}^0$ actually exceeds the proposed upper bound $\overline{\gamma}^0$.
\end{example}

\begin{figure}[ht]
\begin{center}
\input{tikz/directRbad.tex}
\end{center}
\caption{For finest-level box constraints $\underline{\gamma}^1,\overline{\gamma}^1$ like these, direct restriction using any of $\iR,\minR,\maxR$, to a coarser mesh (solid dots), is problematic for admissibility of coarse corrections.}
\label{fig:directRbad}
\end{figure}
\pef{Can we put another figure beside it illustrating the choice $\underline{\gamma}^0, \overline{\gamma}^0 \equiv \maxR \underline{\gamma}^1, \minR \overline{\gamma}^1$?}
\pef{Maybe use different colours for upper and lower bounds?}

@@

This construction of multilevel box constraints, fundamentally the same as the unilateral construction in \cite{GraeserKornhuber2009}, is designed to permit large-yet-admissible coarse corrections, which promotes multilevel solver efficiency.  Note that for Algorithm \ref{alg:fascd} below, any system satisfying \eqref{eq:fe:chiordering} will suffice; the particular formulas in \eqref{eq:fe:chilevels} are not required.

We now compute the LDC differences \cite{GraeserKornhuber2009}:
\pef{Say a few words of motivation first for introducing this concept.}
\begin{equation}
\underline{\phi}^j = \underline{\chi}^j - \underline{\chi}^{j-1}, \qquad \overline{\phi}^j = \overline{\chi}^j - \overline{\chi}^{j-1}.  \label{eq:fe:philevels}
\end{equation}
We take $\underline{\chi}^{-1}=\overline{\chi}^{-1}=0$ so that $\underline{\phi}^0=\underline{\chi}^0$ and $\overline{\phi}^0=\overline{\chi}^0$.  Regarding subtraction of extended real-valued quantities, if $\underline{\chi}^j(x_p)=-\infty$ then we set $\underline{\phi}^j(x_p)=-\infty$ regardless of the value of $\underline{\chi}^{j-1}(x_p)$, and similarly $\overline{\phi}^j(x_p)=+\infty$ if $\overline{\chi}^j(x_p)=+\infty$.  While $\underline{\phi}^{j},\overline{\phi}^{j} \in \tilde{\mathcal{V}}^J$ again bracket zero, $\underline{\phi}^j \le 0 \le \overline{\phi}^j$, they are not ordered as in \eqref{eq:fe:chiordering}.

@@

Our V-cycle solver (Algorithm \ref{alg:fascd}) will compute corrections in $\mathcal{D}^j$ for the downward stage and in $\mathcal{U}^j$ for the upward stage. This strategy is, however, based on \emph{incomplete} CDs, in the general box-constrained cases, as we now explain.  Each upward set $\mathcal{U}^j$ is generally only a superset of a sum of downward sets $\mathcal{D}^j$, but this becomes a true CD, i.e.~satisfying \eqref{eq:subspacedecomp}--\eqref{eq:constraintrestrictionsum}, in unilateral cases.  The following Lemma, apparently new, is based in part on the unilateral CD construction underlying Algorithm 4.7 in \cite{GraeserKornhuber2009}; see also equation (59) in \cite{Tai2003}.

@@

\begin{example}  \label{ex:notfullcd}
Let $a > 0$ be an arbitrary positive number.  Consider a 2-level hierarchy ($J=1$), and suppose the lower defect constraint is the constant function $\underline{\chi}^1=-a$.  Note $\underline{\chi}^0=\maxR \underline{\chi}^1=-a$ also, so $\underline{\phi}^1=0$ and $\underline{\phi}^0=-a$.  Suppose $z^1$ is a ``sawtooth'' function which takes on value $-a$ on the coarse nodes and value $0$ on the other nodes.  Then, for any $\overline{\chi}^1\ge 0$, we have $\overline{\chi}^1 \ge 0 \ge z^1\ge \underline{\chi}^1$, so $z^j \in \mathcal{U}^1$.  However, $I_{1\to 0}^\ominus(z^1 - \underline{\chi}^1) = \minR(z^1 - \underline{\chi}^1) = 0$ identically.  From \eqref{eq:fe:unilateraldecompositionoperator}, $\Pi_1 z^1$ simplifies to $\Pi_1 z^1 = z^1 - \underline{\chi}^1$.  But then $\Pi_1 z^1$ attains a maximum $a\ge 0$.  This maximum can be chosen to exceed any upper defect obstacle $\overline{\phi}^1$, so $\Pi_1 z^1 \notin \mathcal{D}^1$.
\end{example}

@@

Inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion}, which are special to multilevel schemes based on defect constraints, hold regardless of whether a CD operator $\Pi_i$ can be constructed.  Here there is, in fact, another ``basic'' iteration, an incomplete variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}) that does not need the $\Pi_i$ operators.  It is a variant of \pr{cd-mult} (Algorithm \ref{alg:basiccd-mult}).

@@

While inclusions \eqref{eq:fe:downwardsuminclusion} and \eqref{eq:fe:upwardsuminclusion} suffice to show the admissibility of all iterates and corrections in Algorithm \ref{alg:fascd} below, a convergence analysis, perhaps extending the arguments of \cite{Tai2003} or \cite{GraeserKornhuber2009} to VIs which are not of minimization type, and/or to box constraints, would seem to require full CDs.  We cannot currently offer that analysis.

@@

A key idea is that, because of inclusions \eqref{eq:fe:upwardsuminclusion}, up-smoothing solves less-restricted VI problems than down-smoothing.  Intuitively-speaking, going downward one does not know what the next coarse corrections will be, and thus on any level one must choose conservative constraint sets $\mathcal{D}^j$ so that the sum of the corrections is admissible once we return to a given level.  Going upward, however, no further coarse corrections can violate admissibility.
